Using simulation to optimize the sampling design of fisheries-independent trawl surveys

### 2018-07-03 analysis/paper/sim_survey_paper.Rmd
Edit current draft of the paper

---
title: Using simulation to optimize the sampling design of fisheries-independent trawl surveys
author: "Paul M. Regular, Fran Mowbray"
date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
affiliation: "This is a test"
output: 
  word_document:
    reference_docx: template.docx
bibliography: references.bib
csl: ices-journal-of-marine-science.csl
---

@@

A central question to fish stock assessments is whether enough samples were collected to adequately represent the population being monitored. For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling. We therefore need to assess the impact of haul design and length and age sampling on survey accuracy to determine optimal sampling effort. In this study we used simulations to try to answer this question. The method draws repeated stratified-random samples from a spatially correlated field of simulated fish; these samples were then analyzed and estimates of abundance at age were compared against a known truth. By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population. Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent. These results indicate that length and age sampling efforts for cod (*Gadus morhua*) in the trawl survey conducted by Fisheries and Oceans Canada, Newfoundland region, can be scaled back without significant losses to precision. Time saved by reducing said sampling may afford more time to conduct more sets and/or increase sampling efforts of other species. While this simulation is tailored to a specific case, the method may be adapted to different systems and survey procedures.

@@

Two main approaches have been applied to try and answer this question: one involves statistical estimators of effective sample size [@Pennington2002; @Stewart2014], and the other employs resampling methods to test alternate sampling protocol [@Cervino2006; @Zhang2013]. Both approaches have shown that lengths are usually oversampled and a reduction in said sampling can often be done without significant loss of accuracy [@Pennington2002]. The same result holds for other positively correlated characteristics, such as age [@Coggins2013] and stomach contents [@Bogstad1995]. Pennington and Vølstad [-@Pennington1994] concluded that the best way to improve the accuracy of surveys with strong intra-haul correlation is to take fewer samples per sampling unit and increase the total number of sampling locations. Few studies, however, have tested these conclusions using simulations. A key advantage of a pure simulation approach is that "true" population size and all stages of sampling can be controlled. The relative effects of each stage of sampling on our perception of the "truth" can then be tested. This level of testing is of praticular importance for age composition estimates, and the contemporary age-based assessment models that rely on these estimates, as they are affected by all stages of sampling. A key challenge, of course, is building a simulation that emulates reality.

@@

>>>>

---
title: Using simulation to optimize the sampling design of fisheries-independent trawl surveys
author: "Paul M. Regular, Fran Mowbray, et al.?"
date: "Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada"
output: 
  word_document:
    reference_docx: template.docx
bibliography: references.bib
csl: ices-journal-of-marine-science.csl
---

@@

A central question to fish stock assessments is: how many samples are needed to adequately represent the population being monitored? For assessments based on trawl surveys, our ability to estimate population characteristics, such as abundance at age, is affected by multiple levels of sampling (i.e. the number of sets, lengths and ages sampled in a survey). We therefore need to assess the impact of haul design and length and age sampling on survey accuracy to determine optimal sampling effort. In this study we simulated repeated stratified-random samples from a spatially correlated field of simulated fish. These samples were then analyzed and estimates of abundance at age were compared against a known truth. By varying the sampling protocol, we demonstrate that increasing the number of sets conducted per stratum provides the greatest gains in precision when sampling a clustered population. Increasing length and age sampling efforts also improved abundance estimates, but to a lesser extent. These results indicate that length and age sampling efforts for some species(e.g. cod *Gadus morhua*) in the trawl survey conducted by regional fisheries management organizations can be scaled back without significant losses to precision. Time saved by reducing said sampling may afford more time to conduct additional sets and/or increase sampling efforts of other species. While this simulation is tailored to a specific case, the method may be adapted to different systems and survey procedures.

@@

Two main approaches have been applied to try and answer this question: one involves statistical estimators of effective sample size [@Pennington2002; @Stewart2014], and the other employs resampling methods to test alternate sampling protocol [@Cervino2006; @Zhang2013]. Both approaches have shown that lengths are usually oversampled and a reduction in said sampling can often be done without significant loss of accuracy [@Pennington2002]. The same result holds for other positively correlated characteristics, such as age [@Coggins2013] and stomach contents [@Bogstad1995]. Pennington and Vølstad [-@Pennington1994] concluded that the best way to improve the accuracy of surveys with strong intra-haul correlation is to take fewer samples per sampling unit and increase the total number of sampling locations. Few studies, however, have tested these conclusions using simulations. A key advantage of a pure simulation approach is that "true" population size and all stages of sampling can be controlled (set, length and age sampling protocol). The relative effects of each stage of sampling on our perception of the "truth" can then be tested. This level of testing is of particular importance for age composition estimates, and the contemporary age-based assessment models that rely on these estimates, as they are affected by all stages of sampling. A key challenge, of course, is building a simulation that emulates reality.

@@

# Acknowledgements

- Feedback and advice: Alejandro Buren, Mariano Koen-Alonso, Karen Dwyer, Joanne Morgan, Geoff Evans, Don Power, Brian Healey, Martha Robertson, Danny Ings, Don Stansbury, Mark Simpson, Derek Osborne, Peter Upward, Brad Squires, Dwayne Pittman, Paul Higdon
- Dave Cote for a review of a pervious version of this manuscript
